---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo=FALSE, message=FALSE, warning=FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)

library(introDS)
```

[![Travis-CI Build Status](https://travis-ci.org/SMAC-Group/introDS.svg?branch=master)](https://travis-ci.org/SMAC-Group/introDS)
[![Project Status: Active](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![Licence](https://img.shields.io/badge/licence-AGPL--3.0-blue.svg)](https://opensource.org/licenses/AGPL-3.0)
[![minimal R version](https://img.shields.io/badge/R%3E%3D-3.5.0-6666ff.svg)](https://cran.r-project.org/)


# Introduction to Data Science

The objective of the `introDS` package is to provide a support for the course entitled "Introduction to Data Science" given at University of Geneva. This course is intended to provide an introduction to statistical programming using the R language. It will also provide students with notions of data management, manipulation and analysis as well as of reproducible research, result-sharing and version control (using GitHub). At the end of the class, students should be able to automatically extract data from websites and create interactive web apps as well as dynamic reports to visualize and analyze them. Students should also be able to construct their own R packages, make them available on GitHub and document them using literate programming.

# Installation Instructions

The `introDS` package is available on both CRAN and GitHub. The CRAN version is considered stable while the GitHub version is subject to modifications/updates which may lead to installation problems or broken functions. You can install the stable version of the `introDS` package with:

```{r, eval = F}
install.packages("introDS")
```

For users who are interested in having the latest developments, the GitHub version is ideal although more dependencies are required to run a stable version of the package. 

```{r, eval = F}
# Install dependencies
install.packages(c("devtools","knitr","rmarkdown"))

# Install the package from GitHub
devtools::install_github("SMAC-Group/introDS")
```

